---
title: "Penguin Assignment"
author: "Anonymous"
date: "04/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Organising and cleaning data:

Setting working directory as "PenguinProjects" folder (**please adjust to suit your working directory**):

```{r}
setwd("~/RWorkingDirectory/PenguinProjects")
```

Loading in libraries, data and functions:

```{r}
source("functions/libraries.r")
source("functions/cleaning.r")
source("functions/plotting.r")
```

Creating a new folder for the raw data and saving the raw data:

```{r}
#Creating a new data folder
dir.create("data")

#Creating a new raw data folder
dir.create("data/data_raw")

#Saving the raw data
write.csv(penguins_raw, "data/data_raw/penguins_raw.csv")
```

Cleaning the data, creating a new folder for the clean data and saving the clean data:

```{r}
#Using function from "cleaning.r" to clean the data
penguins_clean <- cleaning(penguins_raw)

#Creating a new clean data folder
dir.create("data/data_clean")

#Saving the clean data
write.csv(penguins_clean, "data/data_clean/penguins_clean_05-12-2022.csv")
```

Looking at the data:

```{r}
head(penguins_clean)
```



### Preparing data for analysis:

Removing rows without sex or body mass recorded:

```{r}
#Using function from "cleaning.r" to prepare data
penguins_mass <- remove_empty_body_mass(penguins_clean)

#Saving mass data
write.csv(penguins_mass, "data/data_clean/penguins_mass_05-12-2022.csv")

#Displaying the data
head(penguins_mass)
```

Defining factors and numerical variables:

```{r, eval = FALSE}
penguins_mass$species <- as.factor(penguins_mass$species)
penguins_mass$sex <- as.factor(penguins_mass$sex)
penguins_mass$body_mass_g <- as.numeric(penguins_mass$body_mass_g)
```



### Analysing the data:

**Q: To what extent does the species and sex of a penguin influence body mass?**

Creating a model with no interaction and testing assumptions of statistical analysis:

```{r}
#Creating the model
lmSpecies_Sex_Mass_No_Interaction <- aov(body_mass_g ~ species + sex, data = penguins_mass)

#Plotting to test assumptions
par(mfrow = c(2,2))
plot(lmSpecies_Sex_Mass_No_Interaction)
```
Overall, the data appears to meet assumptions of ANOVA.

Running ANOVA:

```{r}
summary(lmSpecies_Sex_Mass_No_Interaction)
```
Both species and sex appear to have a significant effect on body mass.

Building a model that takes into account interaction between species and sex:

```{r}
#Creating the model
lmSpecies_Sex_Mass_Interaction <- aov(body_mass_g ~ species * sex, data = penguins_mass)

#Plotting to test assumptions
par(mfrow = c(2,2))
plot(lmSpecies_Sex_Mass_Interaction)
```
Overall, the data appears to fit the assumptions of ANCOVA.

Running ANCOVA:

```{r}
summary(lmSpecies_Sex_Mass_Interaction)
```
There is a significant effect of interaction between species and sex on body mass.

Comparing the two models:

```{r}
anova(lmSpecies_Sex_Mass_No_Interaction,lmSpecies_Sex_Mass_Interaction)
```
Adding an interaction between species and sex significantly improves the model.

Running a Tukey HSD test:

```{r, eval = FALSE}
#Using TukeyHSD test to compare between groups (results hidden here to keep document tidy)
TukeyHSD(lmSpecies_Sex_Mass_Interaction)
```
The Tukey HSD test suggests significant differences between body mass of penguins grouped by species and sex.
The comparisons between groups which are not significantly different are:
1) Chinstrap females and Adelie females
2) Chinstrap males and Adelie males



### Producing figures to display the data:

Creating a new folder to store figures in:

```{r}
dir.create("figures")
```

Producing a dot plot with lines for each species:

```{r}
source("functions/plotting.r")
#Using function from "plotting.R" to produce graph
plot_mass_figure(penguins_mass)
```

Saving the figure as a PNG file:

```{r}
#Using function from "plotting.R" to save figure as PNG
save_mass_plot_png(penguins_mass, "figures/fig_01_06-12-2022.png", 20, 600, 1.4)
```

Saving the figure as an SVG file:

```{r}
#Using function from "plotting.R" to save figure as SVG
save_mass_plot_svg(penguins_mass, "figures/fig_01_06-12-2022.svg", 20, 1.4)
```
